---
title: "Backup nbarapm.com API Data"
format: html
---

## Overview

This document fetches the complete raw response from each of the four
nbarapm.com JSON endpoints and saves them as CSV files in the
`nbarapm_backup/` folder.

Run this periodically (or whenever you update the main app data) to keep a
local snapshot in case the API becomes unavailable or requires authentication.
Each run overwrites the previous backup.

The backup files are **unfiltered and unmodified** — every column and every
season that the API returns is preserved so nothing is lost.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(jsonlite)

BACKUP_DIR <- file.path(dirname(rstudioapi::getSourceEditorContext()$path),
                        "nbarapm_backup")
dir.create(BACKUP_DIR, showWarnings = FALSE)
```

```{r helpers}
fetch_and_save <- function(url, filename_base) {
  cat("Fetching", url, "...\n")
  tryCatch({
    raw <- fromJSON(url, simplifyDataFrame = TRUE) |> as_tibble()
    cat("  -> ", nrow(raw), "rows,", ncol(raw), "columns\n")

    # Normalise hyphenated column names for CSV safety (O-MAMBA -> O_MAMBA)
    raw <- raw |> rename_with(~ str_replace_all(.x, "-", "_"))

    out_path <- file.path(BACKUP_DIR, paste0(filename_base, ".csv"))
    write_csv(raw, out_path)
    cat("  Saved:", out_path, "\n\n")

    raw
  }, error = function(e) {
    warning("Failed to fetch ", url, ": ", conditionMessage(e))
    NULL
  })
}
```

```{r fetch-mamba}
mamba_raw <- fetch_and_save("https://www.nbarapm.com/load/mamba",  "mamba")
```

```{r fetch-raptor}
raptor_raw <- fetch_and_save("https://www.nbarapm.com/load/raptor", "raptor")
```

```{r fetch-darko}
darko_raw <- fetch_and_save("https://www.nbarapm.com/load/DARKO",  "darko")
```

```{r fetch-lebron}
lebron_raw <- fetch_and_save("https://www.nbarapm.com/load/lebron", "lebron")
```

```{r summary}
datasets <- list(MAMBA = mamba_raw, RAPTOR = raptor_raw,
                 DARKO = darko_raw, LEBRON = lebron_raw)

cat("=== Backup summary ===\n")
for (nm in names(datasets)) {
  d <- datasets[[nm]]
  if (is.null(d)) {
    cat(nm, ": FAILED\n")
  } else {
    yr_col <- if ("year" %in% names(d)) "year" else "season"
    yr_range <- paste(range(d[[yr_col]], na.rm = TRUE), collapse = " - ")
    cat(nm, ":", nrow(d), "rows |", ncol(d), "cols | years", yr_range, "\n")
  }
}

cat("\nFiles written to:", normalizePath(BACKUP_DIR), "\n")
```
