---
title: "Scrape NBA Advanced Stats by Team"
format: html
editor: visual
---

```{r setup}
# install.packages("polite")  # uncomment if needed
library(polite)
library(rvest)
library(tidyverse)
library(janitor)
library(glue)
```

## Overview

Scrapes the **Advanced** stats table from each NBA team page on Basketball Reference (`/teams/{TEAM}/{YEAR}.html`) for every season we have other scraped data for (2016–2025, where the year is the final year of the season — e.g. 2024 = 2023-24).

**Columns extracted:** Player, Age, Pos, G, MP, PER, OWS, DWS, WS, OBPM, DBPM, BPM, VORP.

> **Note:** Stats are per team stint. A player traded mid-season will appear once per team they played for, not as a single combined row. The `team` column identifies which stint each row belongs to. Deduplicate or sum as needed for full-season totals.

This makes **300 requests** (30 teams × 10 seasons). With Basketball Reference's 3-second crawl delay, expect \~15 minutes of run time.

------------------------------------------------------------------------

## Configuration

```{r config}
# Seasons to scrape: year = final year of the season (2024 = 2023-24)
seasons <- 2016:2025

# All 30 NBA team abbreviations (stable since 2016)
nba_teams <- c(
  "ATL", "BOS", "BRK", "CHO", "CHI", "CLE", "DAL", "DEN", "DET", "GSW",
  "HOU", "IND", "LAC", "LAL", "MEM", "MIA", "MIL", "MIN", "NOP", "NYK",
  "OKC", "ORL", "PHI", "PHO", "POR", "SAC", "SAS", "TOR", "UTA", "WAS"
)

# Columns to keep from the advanced table (exact BBRef names)
KEEP_COLS <- c("Player", "Age", "Pos", "G", "MP",
               "PER", "OWS", "DWS", "WS", "OBPM", "DBPM", "BPM", "VORP")

OUTPUT_PATH <- "/Users/willbrett/Documents/Basketball Project/advanced_stats.csv"
```

## Polite session

```{r session}
bbref_session <- bow(
  "https://www.basketball-reference.com",
  user_agent = "Basketball Research Project (academic use)"
)
```

## Parsing helper

BBRef wraps many tables in HTML comments. We strip those before parsing, then use the first row as column names and filter out non-player rows.

```{r helper}
parse_advanced_table <- function(page, team, year) {
  # Strip HTML comments so buried tables become visible
  page_text <- gsub("<!--|-->", "", as.character(page))
  page2     <- read_html(page_text)

  node <- page2 |> html_node("#advanced")
  if (inherits(node, "xml_missing")) return(NULL)

  raw      <- node |> html_table(header = FALSE)
  col_names <- as.character(raw[1, ])   # row 1 is the real header
  df        <- raw[-1, ]
  names(df) <- col_names

  df |>
    filter(
      Player != "Player",   # drop repeated header rows
      Player != "",
      !str_detect(Player, "^Team|^League")  # drop totals rows
    ) |>
    select(any_of(KEEP_COLS)) |>
    mutate(
      team   = team,
      season = year,
      .before = 1
    ) |>
    mutate(across(any_of(c("Age","G","MP","PER","OWS","DWS","WS",
                            "OBPM","DBPM","BPM","VORP")),
                  ~ suppressWarnings(as.numeric(.x))))
}
```

## Scrape all teams and seasons

```{r scrape}
all_advanced <- map_dfr(seasons, function(yr) {
  message("\n── Season ", yr, " ─────────────────────────────────────")

  map_dfr(nba_teams, function(tm) {
    path   <- glue("/teams/{tm}/{yr}.html")
    result <- tryCatch({
      page <- nod(bbref_session, path) |>
        scrape(content = "text/html; charset=UTF-8")
      parse_advanced_table(page, tm, yr)
    }, error = function(e) {
      warning(tm, "/", yr, ": ", conditionMessage(e))
      NULL
    })

    if (!is.null(result)) {
      message("  ", tm, " — ", nrow(result), " players")
    }
    result
  })
})

message("\nDone. Total rows: ", nrow(all_advanced))
```

## Inspect

```{r inspect}
glimpse(all_advanced)

# Season coverage
all_advanced |>
  count(season, name = "player_rows") |>
  print()

# Quick sanity check: top PER players in 2024
all_advanced |>
  filter(season == 2024) |>
  slice_max(PER, n = 10) |>
  select(season, team, Player, Pos, G, PER, WS, BPM, VORP) |>
  print()
```

## Save

```{r save}
write_csv(all_advanced, OUTPUT_PATH)
message("Saved ", nrow(all_advanced), " rows to advanced_stats.csv")
```
