---
title: "Scrape NBA Awards from Basketball Reference"
format: html
editor: visual
---

```{r setup}
# install.packages("polite")  # uncomment if not yet installed
library(polite)
library(rvest)
library(tidyverse)
library(janitor)
library(glue)
```

## Seasons

We have stint-level data for the 2016–2024 seasons, matching the processed play-by-play RDS files.

```{r seasons}
seasons <- 2016:2025
```

## Polite session

`bow()` reads `robots.txt` and respects the declared crawl delay. All subsequent requests go through `nod()` + `scrape()`, which enforces rate-limiting automatically.

```{r session}
bbref_session <- bow(
  "https://www.basketball-reference.com",
  user_agent = "Basketball Research Project (academic use)"
)
```

## Scraping helpers

Basketball Reference's award tables (`/awards/awards_{year}.html`) have a two-row header and occasional empty separator rows. We read with `header = FALSE`, promote row 2 to column names, then filter out non-player rows.

Table IDs (confirmed from HTML inspection):

| Award               | Table `id`            |
|---------------------|-----------------------|
| MVP voting          | `mvp`                 |
| All-NBA Teams       | `leading_all_nba`     |
| All-Defensive Teams | `leading_all_defense` |

```{r helpers}
# Read a BBRef award table robustly:
#   - use row 2 as the header (skipping the group-label row)
#   - drop empty / repeated-header rows
#   - clean column names with janitor
parse_bbref_award_table <- function(page, css_id) {
  node <- page |> html_node(glue("#{css_id}"))
  if (inherits(node, "xml_missing")) return(NULL)

  raw <- node |> html_table(header = FALSE)

  # Row 2 contains the real column labels
  col_names <- as.character(raw[2, ])
  df <- raw[-(1:2), ]            # drop both header rows
  names(df) <- col_names

  df |>
    filter(Player != "", Player != "Player") |>  # drop empties & repeated headers
    clean_names() |>
    # Keep everything as character so seasons bind cleanly; convert numerics below
    mutate(across(everything(), as.character))
}
```

## Per-season scrape function

```{r scrape_fn}
scrape_season_awards <- function(year, session) {
  path <- glue("/awards/awards_{year}.html")
  page <- nod(session, path) |> scrape(content = "text/html; charset=UTF-8")

  mvp        <- parse_bbref_award_table(page, "mvp")
  all_nba    <- parse_bbref_award_table(page, "leading_all_nba")
  all_defense <- parse_bbref_award_table(page, "leading_all_defense")

  list(
    mvp         = if (!is.null(mvp))         mutate(mvp,         season = as.integer(year), .before = 1),
    all_nba     = if (!is.null(all_nba))     mutate(all_nba,     season = as.integer(year), .before = 1),
    all_defense = if (!is.null(all_defense)) mutate(all_defense, season = as.integer(year), .before = 1)
  )
}
```

## Scrape all seasons

`polite` enforces the crawl delay between requests automatically; no extra `Sys.sleep()` needed.

```{r scrape_all}
awards_list <- map(seasons, function(yr) {
  message("Scraping ", yr, "...")
  tryCatch(
    scrape_season_awards(yr, bbref_session),
    error = function(e) {
      warning("Failed for season ", yr, ": ", conditionMessage(e))
      NULL
    }
  )
})

names(awards_list) <- seasons
```

## Combine into tidy data frames

```{r combine}
awards_list <- compact(awards_list)

mvp_voting  <- map_dfr(awards_list, "mvp")
all_nba     <- map_dfr(awards_list, "all_nba")
all_defense <- map_dfr(awards_list, "all_defense")
```

## Convert numeric columns

```{r coerce_numeric}
# Columns that should be numeric across all three tables
mvp_num_cols     <- c("rank", "age", "first", "pts_won", "pts_max", "share",
                       "g", "mp", "pts", "trb", "ast", "stl", "blk",
                       "fg_percent", "x3p_percent", "ft_percent", "ws", "ws_48")
all_nba_num_cols <- c("age", "pts_won", "pts_max", "share",
                       "x1st_tm", "x2nd_tm", "x3rd_tm",
                       "g", "mp", "pts", "trb", "ast", "stl", "blk",
                       "fg_percent", "x3p_percent", "ft_percent", "ws", "ws_48")
def_num_cols     <- c("age", "pts_won", "pts_max", "share",
                       "g", "mp", "pts", "trb", "ast", "stl", "blk",
                       "fg_percent", "x3p_percent", "ft_percent",
                       "ws", "ws_48", "dws", "dbpm", "drtg")

coerce_numeric <- function(df, num_cols) {
  cols_present <- intersect(num_cols, names(df))
  df |> mutate(across(all_of(cols_present), ~ suppressWarnings(as.numeric(.x))))
}

mvp_voting  <- coerce_numeric(mvp_voting,  mvp_num_cols)
all_nba     <- coerce_numeric(all_nba,     all_nba_num_cols)
all_defense <- coerce_numeric(all_defense, def_num_cols)
```

## Clean up key columns

The `number_tm` column encodes the team tier: - All-NBA: `1T` / `2T` / `3T` (teams), `ORV` (other receiving votes) - All-Defensive: `1st` / `2nd`, `ORV`

```{r clean}
# Rename the team-tier column to something clearer
all_nba <- all_nba |>
  rename(team_tier = number_tm) |>
  mutate(
    nba_team = case_when(
      team_tier == "1T"  ~ "1st Team",
      team_tier == "2T"  ~ "2nd Team",
      team_tier == "3T"  ~ "3rd Team",
      team_tier == "ORV" ~ "Other Votes",
      TRUE               ~ team_tier
    )
  )

all_defense <- all_defense |>
  rename(team_tier = number_tm) |>
  mutate(
    defense_team = case_when(
      team_tier == "1st" ~ "1st Team",
      team_tier == "2nd" ~ "2nd Team",
      team_tier == "ORV" ~ "Other Votes",
      TRUE               ~ team_tier
    )
  )
```

## Inspect

```{r inspect}
# ── MVP winners ──────────────────────────────────────────────────────────────
cat("=== MVP Winners by Season ===\n")
mvp_voting |>
  filter(rank == 1) |>
  select(season, player, tm, first, pts_won, pts_max, share) |>
  print(n = Inf)

# ── All-NBA 1st Team ─────────────────────────────────────────────────────────
cat("\n=== All-NBA 1st Team by Season ===\n")
all_nba |>
  filter(nba_team == "1st Team") |>
  select(season, pos, player, tm) |>
  arrange(season, pos) |>
  print(n = Inf)

# ── All-Defensive 1st Team ───────────────────────────────────────────────────
cat("\n=== All-Defensive 1st Team by Season ===\n")
all_defense |>
  filter(defense_team == "1st Team") |>
  select(season, pos, player, tm) |>
  arrange(season, pos) |>
  print(n = Inf)
```

## Save outputs

```{r save}
output_dir <- "/Users/willbrett/Documents/Basketball Project"

write_csv(mvp_voting,  file.path(output_dir, "awards_mvp_voting.csv"))
write_csv(all_nba,     file.path(output_dir, "awards_all_nba.csv"))
write_csv(all_defense, file.path(output_dir, "awards_all_defense.csv"))

message(
  "Saved:\n",
  "  awards_mvp_voting.csv  (", nrow(mvp_voting),  " rows)\n",
  "  awards_all_nba.csv     (", nrow(all_nba),     " rows)\n",
  "  awards_all_defense.csv (", nrow(all_defense), " rows)"
)
```
